---
title: "Getting started with slashOhdsiAcpClient"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with slashOhdsiAcpClient}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`slashOhdsiAcpClient` is a low-level R client for a running OHDSI StudyAgent
ACP bridge. Prefer its explicit-client API: construct a client once, then pass
it to the flow or action wrapper you need.

## Configure and connect

Use environment variables for deployment-specific settings. Keep tokens outside
R source files.

```{r connect, eval = FALSE}
library(slashOhdsiAcpClient)

acp_url <- Sys.getenv("ACP_URL", "http://127.0.0.1:8765")
acp_token <- Sys.getenv("ACP_TOKEN", unset = "")
Sys.setenv(ACP_TIMEOUT = "1800")

client <- acp_client(
  url = acp_url,
  token = if (nzchar(acp_token)) acp_token else NULL,
  check = TRUE
)

acp_server_info(client)
```

By default, `acp_client(check = TRUE)` calls `/health` and verifies that the
bridge reports a compatible ACP API version. `ACP_TIMEOUT` is measured in
seconds and defaults to 180.

## Request a flow

Call the typed wrapper rather than constructing endpoint paths manually. For
example, request phenotype recommendations with a clear study intent:

```{r phenotype-recommendation, eval = FALSE}
recommendations <- acp_suggest_phenotypes(
  client = client,
  study_intent = paste(
    "Estimate the comparative effectiveness of first-line treatments for",
    "adults with type 2 diabetes, using time to hospitalization as an outcome."
  ),
  top_k = 20,
  max_results = 3,
  candidate_limit = 10
)

str(recommendations, max.level = 2)
```

Responses are parsed R lists. Their contents are determined by the ACP service,
so inspect them before assuming optional fields exist.

## Local artifact flows

Some flows accept paths to local protocol, cohort, or concept-set artifacts.
The client validates and reads those paths locally, then sends the resulting
content to ACP. For example:

```{r local-artifacts, eval = FALSE}
review <- acp_review_phenotypes(
  client = client,
  protocol_path = "protocol.md",
  cohort_paths = c("cohorts/type_2_diabetes.json")
)
```

Use paths only for files the current R process can read. The ACP bridge receives
the file contents; it does not need to share the client's file system.

## Default-client compatibility helpers

`acp_connect()` stores a checked default client for compatibility-oriented,
interactive helpers. New reusable code should normally keep passing an explicit
`client`.

```{r default-client, eval = FALSE}
acp_connect(url = acp_url, token = if (nzchar(acp_token)) acp_token else NULL)

suggestions <- suggestPhenotypes(
  studyIntent = "Identify suitable cohorts for a new-user study of statins.",
  interactive = FALSE
)
```

When no default client has been connected, certain compatibility helpers return
a local stub result whose `source` is `"stub_no_acp"`; that is not an ACP
response.

## Troubleshooting

| Symptom | Meaning and action |
|---|---|
| `ACP bridge not reachable` | The client could not obtain a successful response from `/health`. Start the bridge and verify `ACP_URL`, network access, and any bridge authentication requirement. |
| Incompatible ACP API version | Upgrade or align the R package and StudyAgent ACP bridge to a compatible API version. |
| `stub_no_acp` response | A compatibility helper was called without a default connection. Call `acp_connect()`, or use an explicit client with a typed wrapper. |
| Missing local JSON or protocol file | Correct the supplied path and ensure the R process can read it. |
| `ACP error` containing an LLM or embedding HTTP 401 | The request reached ACP, but ACP could not authenticate to its configured LLM or embedding provider. This is different from an ACP bearer-token error. Ask the ACP deployment administrator to configure or renew the provider credentials in the ACP environment. |

For example, a phenotype request may fail with an error resembling:

```text
ACP error: {"status":"error","error":"phenotype_search_failed",
"details":{"details":"Embedding request failed: HTTP Error 401: Unauthorized",
"_meta":{"tool":"phenotype_search"}}}
```

The `Embedding request failed` and `phenotype_search` details identify an
ACP-side provider authorization problem. Changing `ACP_TOKEN` in the R client
will not fix it unless the bridge itself also rejected that token. When seeking
support, provide the flow name and sanitized error payload, but never include
tokens or provider secrets.
